---
title: "CarrotTransformation"
author: "Pankaj Bhatta"
date: "04/05/2021"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Necessary items:
- **libraries:** tidyverse, ggsci, ggforce, patchwork, Hmisc
- **files:** carrot_df.csv

Within RStudio you can directly edit these blocks of code and turn in your homework that way with your code and figures printed.   
I **HIGHLY** recommend getting your code to work in a normal R script then just copy and pasting the final code over to this document

**First: import libraries, set your working directory, and read in bloom_df** 

```{r init, message=FALSE, warning=FALSE}
library(tidyverse) 
library(ggforce) 
library(ggsci)
library(patchwork)
library(Hmisc)
setwd('~/Desktop/BIOL792_2') #change to match your ggplot directory
carrot_df <- read.csv('carrot_transformation_version2-1.csv')
```
### carrot_df contents
- **carrot_variety**
- **carrot_parts** 
- **agrobacteria_type**  
- **gene**  
- **conditions**  
- **time_in_MS1D_dark**  
- **time_in_MS1D_dl**  
- **time_in_half_MS1D_dl**  
- **time_in_quarter_MS1D_dl**  
- **antibiotics_in_MS1D_dl**  
- **antibiotics_in_half_MS1D_dl**  
- **antibiotics_in_qaurter_MS1D_dl**  
- **callus_status**  
- **PCR_Gel_status**  
- **GC_MS_polyacetylene_percentage** 

Let's take a peak and look at the structure

```{r carrot_df}
carrot_df[sample(nrow(carrot_df), 5), ]
summary(carrot_df)
```
##### We are creating ggplot of the following data for the visualization.
##### The layout will be: 
- x axis: carrot_variety/carrot_parts/agrobacteria_type/gene  
- y axis: value of GC_MS_polyacetylene_percentage 

##### We are creating two plots for visualization:  
- bar and error bars (mean and 95% conf. int.)  
- raw data + point and error bars (mean and 95% conf. int.)  
```{r carrot_long_df}
carrot_long_df <- carrot_df %>%
  gather(key=polyacetylene, value = mean, c(GC_MS_polyacetylene_percentage))
```

#### Calculation and summarization of polyacetylene based on carrot_variety

```{r 1.1}
variety_polyacetylene_sum_df <- carrot_df %>%
  group_by(carrot_variety) %>%
  summarise(mean = mean(GC_MS_polyacetylene_percentage, na.rm= TRUE),
            sd = sd(GC_MS_polyacetylene_percentage, na.rm= TRUE),
            n = n()) %>%
  mutate(se = sd / sqrt(n),
         ci = 1.96*se)
variety_polyacetylene_sum_df
```
#### Bar and error bars (mean and 95% conf. int.)

```{r 1.2}
carrot_variety_bar <- ggplot(data=variety_polyacetylene_sum_df, aes(x=carrot_variety, y=mean, fill= carrot_variety))+
  geom_bar(stat = 'identity')+
  geom_errorbar(aes(ymin = mean - ci, ymax = mean + ci))+
  xlab('Carrot variety')+
  ylab('Polyacetylene mean')+
  ggtitle('Polyacetylene mean on carrot variety')
  
carrot_variety_bar
```
#### Raw data + point and error bars (mean and 95% conf. int.)

```{r 1.3}
carrot_variety_plots = ggplot(data=carrot_long_df, aes(x=carrot_variety, y=mean, colour=carrot_variety))+
  geom_point(size=1)+
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", color='black')+
  stat_summary(fun = mean, geom = "point", size=3, color='black')+
  xlab('Carrot variety')+
  ylab('Polyacetylene')+
  ggtitle('Polyacetylene vs carrot variety')

carrot_variety_plots
```

####calculation and summarization of polyacetylene based on carrot_parts

```{r 2.1}
parts_polyacetylene_sum_df <- carrot_df %>%
  group_by(carrot_parts) %>%
  summarise(mean = mean(GC_MS_polyacetylene_percentage, na.rm= TRUE),
            sd = sd(GC_MS_polyacetylene_percentage, na.rm= TRUE),
            n = n()) %>%
  mutate(se = sd / sqrt(n),
         ci = 1.96*se)
parts_polyacetylene_sum_df
```

#### Bar and error bars (mean and 95% conf. int.)

```{r 2.2}
carrot_parts_bar <- ggplot(data=parts_polyacetylene_sum_df, aes(x=carrot_parts, y=mean, fill= carrot_parts))+
  geom_bar(stat = 'identity')+
  geom_errorbar(aes(ymin = mean - ci, ymax = mean + ci))+
  xlab('Carrot parts')+
  ylab('Polyacetylene mean')+
  ggtitle('Polyacetylene mean on carrot parts')
carrot_parts_bar
```

#### Raw data + point and error bars (mean and 95% conf. int.)

```{r 2.3}
carrot_parts_plots = ggplot(data=carrot_long_df, aes(x=carrot_parts, y=mean, colour=carrot_parts))+
  geom_point(size=1)+
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", color='black')+
  stat_summary(fun = mean, geom = "point", size=3, color='black')+
  xlab('Carrot parts')+
  ylab('Polyacetylene')+
  ggtitle('Polyacetylene vs carrot parts')

carrot_parts_plots
```

#### Calculation and summarization of polyacetylene based on agrobacteria_type

```{r 3.1}
agrobacteria_polyacetylene_sum_df <- carrot_df %>%
  group_by(agrobacteria_type) %>%
  summarise(mean = mean(GC_MS_polyacetylene_percentage, na.rm= TRUE),
            sd = sd(GC_MS_polyacetylene_percentage, na.rm= TRUE),
            n = n()) %>%
  mutate(se = sd / sqrt(n),
         ci = 1.96*se)
agrobacteria_polyacetylene_sum_df
```
#### Bar and error bars (mean and 95% conf. int.)

```{r 3.2}
agrobacteria_bar <- ggplot(data=agrobacteria_polyacetylene_sum_df, aes(x=agrobacteria_type, y=mean, fill= agrobacteria_type))+
  geom_bar(stat = 'identity')+
  geom_errorbar(aes(ymin = mean - ci, ymax = mean + ci))+
  xlab('Agrobacteria type')+
  ylab('Polyacetylene mean')+
  ggtitle('Polyacetylene mean on Agrobacteria type')
  
agrobacteria_bar
```
#### Raw data + point and error bars (mean and 95% conf. int.)

```{r 3.3}
agrobacteria_plots = ggplot(data=carrot_long_df, aes(x=agrobacteria_type, y=mean, colour=agrobacteria_type))+
  geom_point(size=1)+
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", color='black')+
  stat_summary(fun = mean, geom = "point", size=3, color='black')+
  xlab('Agrobacteria type')+
  ylab('Polyacetylene mean')+
  ggtitle('Polyacetylene vs Agrobacteria type')

agrobacteria_plots
```

#### Calculation and summarization of polyacetylene based on gene

```{r 4.1}
gene_polyacetylene_sum_df <- carrot_df %>%
  group_by(gene) %>%
  summarise(mean = mean(GC_MS_polyacetylene_percentage, na.rm= TRUE),
            sd = sd(GC_MS_polyacetylene_percentage, na.rm= TRUE),
            n = n()) %>%
  mutate(se = sd / sqrt(n),
         ci = 1.96*se)
gene_polyacetylene_sum_df
```
#### Bar and error bars (mean and 95% conf. int.)

```{r 4.2}
gene_bar <- ggplot(data=gene_polyacetylene_sum_df, aes(x=gene, y=mean, fill= gene))+
  geom_bar(stat = 'identity')+
  geom_errorbar(aes(ymin = mean - ci, ymax = mean + ci))+
  xlab('Gene')+
  ylab('Polyacetylene mean')+
  ggtitle('Polyacetylene mean on Gene')
  
gene_bar
```
#### Raw data + point and error bars (mean and 95% conf. int.)

```{r 4.3}
gene_plots = ggplot(data=carrot_long_df, aes(x=gene, y=mean, colour=gene))+
  geom_point(size=1)+
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", color='black')+
  stat_summary(fun = mean, geom = "point", size=3, color='black')+
  xlab('Gene')+
  ylab('Polyacetylene mean')+
  ggtitle('Polyacetylene vs Gene')

gene_plots
```

#### Summarizing bar plots

```{r 5}
carrot_variety_bar + carrot_parts_bar + agrobacteria_bar + gene_bar
```

#### Summarizing point plots

```{r 6}
carrot_variety_plots + carrot_parts_plots + agrobacteria_plots + gene_plots
```

#### Linear regression model

```{r}
smp_siz = floor(0.75*nrow(carrot_df))
set.seed(123)
train_ind = sample(seq_len(nrow(carrot_df)),size = smp_siz)
train = carrot_df[train_ind,]
test = carrot_df[-train_ind,]
linear_model = lm(carrot_df$GC_MS_polyacetylene_percentage ~ carrot_df$carrot_variety + carrot_df$carrot_parts + carrot_df$agrobacteria_type + carrot_df$gene, data = train)
summary(linear_model)
```

#### Prediction in R using linear regression

```{r}
precentage_prediction = predict(linear_model, newdata = test)
actuals_preds <- data.frame(cbind(actuals=carrot_df$GC_MS_polyacetylene_percentage, predicteds=precentage_prediction))
actuals_preds
```




